Interval Probability for Fuzzy Quantum Theories

### 2016-05-17 prob.tex
prob.tex small updates. Need to check the quantum independence...

The conventional definition of a probability
space~\cite{Kolmogorov1950,rohatgi2011introduction,Swart2013}
builds upon the real numbers. In more detail, a probability space
consists of a \emph{sample space} $\Omega$, a space of
\emph{events}~$\events$, and a \emph{probability measure}~$\pmeas$
mapping events in $\events$ to the real interval $[0,1]$. In this
paper, we will only consider \emph{finite} sets of events: we
therefore restrict our attention to non-empty finite sets $\Omega$ as
the sample space. The space of events $\events$ includes every
possible subset of $\Omega$: it is the powerset~$2^\Omega$. Given the
set of events $\events$, a \emph{probability measure} is a function
$\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection $E_{i}$ of pairwise disjoint events, $\pmeas(\bigcup_{i}E_{i})=\tensor*[_{\mathbb{R}}]{\sum}{_{i}}\pmeas(E_{i})$,
where $\tensor*[_{\mathbb{R}}]{\sum}{_{i}}a_{i}$
explicitly specifies $a_{i}\in\mathbb{R}$. We may drop the preposing
$\mathbb{R}$ when there is no confusion. 
\end{itemize}

@@

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware computational or experimental setting,
it is more appropriate to consider probability measures that map
events to a set of elements computable with a fixed set of
resources. We expand on this observation and then consider
%% two approaches from the literature: set-valued probability
%% measures~\cite{Artstein1972,PuriRalescu1983} and 
interval-valued probability
measures~\cite{Dempster1967,Zadeh1986,Weichselberger2000,JamisonLodwick2004}
in detail.\footnote{There is another possible approach that can be
  used to split the real interval $[0,1]$ into a collection of
  subsets~\cite{Artstein1972,PuriRalescu1983} \amr{need to explain the
    connection and why we are not using it.}}

@@

\begin{definition}[Projection Operators;
Orthogonality~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Griffiths2003,Swart2013}]
\label{def:Projection} Given a Hilbert space $\Hilb$, an event\footnote{An event
is formally called an experimental proposition~\cite{BirkhoffVonNeumann1936}, a question~\cite{10.2307/2308516,DBLP:journals/corr/abs-0910-2393}, or an
elementary quantum test~\cite{peres1995quantum}.} mathematically is represented as a projection operator $P:\Hilb\rightarrow\Hilb$ onto a linear subspace~$S$ of $\Hilb$. The set of all events can be defined recursively as follow:
\footnote {``Projection'' is sometimes called
``orthogonal projection" or ``self-adjoint projection" to
emphasize $P^{\dagger} = P$~\cite{Maassen2010}.}
\begin{itemize}
\item $\mathbb{0}$ is a projection. 
\item For any pure state~$\ket{\psi}$, $\proj{\psi}$ is a projection
operator. 
\item Projection operators $P_1$ and $P_2$ are orthogonal if $P_1P_2 =
  P_2P_1 = \mathbb{0}$. The sum of two orthogonal projections~$P_{1}\tensor*[_{\mathscr{O}}]{+}{}P_{2}$
is also a projection, where the preposing subscript $\mathscr{O}$
means $\tensor*[_{\mathscr{O}}]{+}{}$ is an operation between operators,
and may be dropped when there is no ambiguity. \\
\yutsung{Need to prove: \\
The sum of two projection operators~$P_{1}\tensor*[_{\mathscr{O}}]{+}{}P_{2}$
is also a projection operator if and only if they are orthogonal.}
\item Conversely, every projection~$P$ can be expressed as
$\tensor*[_{\mathscr{O}}]{\sum}{_j}\proj{\psi_{j}}$,
where $P$ actually projects onto the linear subspace~$S$ which
has an orthonormal basis~$\{\ket{\psi_{j}}\}$. 
\end{itemize}
\qed\end{definition}

@@

\begin{definition}[Complement; Commutativity;
Independence~\cite{peres1995quantum,Griffiths2003}]~
\begin{itemize}
\item If $P$ is a projection operator, then $\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P$
is also a projection operator. It is orthogonal to $P$, $\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)$,
and it corresponds to the complement event~$\Omega\backslash E$
in classical probability.
\item Projection operators $P_{1}$ and $P_{2}$ commute if $P_{1}P_{2}=P_{2}P_{1}$.
If the projections $P_{1}$ and $P_{2}$ commute, then $P_{1}P_{2}$
is also a projection. This corresponds to the classical intersection
between event. Given a quantum probability measure $\mu$, two projections
$P_{1}$ and $P_{2}$ are independent if $P_{1}$ and $P_{2}$ commute
and $\mu\left(P_{1}P_{2}\right)=\mu\left(P_{1}\right)\mu\left(P_{2}\right)$.
\end{itemize}
\qed\end{definition}

>>>>

The conventional definition of a probability
space~\cite{Kolmogorov1950,Shafer1976,Swart2013}
builds upon the real numbers. In more detail, a probability space
consists of a \emph{sample space} $\Omega$, a space of
\emph{events}~$\events$, and a \emph{probability measure}~$\pmeas$
mapping events in $\events$ to the real interval $[0,1]$. In this
paper, we will only consider \emph{finite} sets of events: we
therefore restrict our attention to non-empty finite sets $\Omega$ as
the sample space. The space of events $\events$ includes every
possible subset of $\Omega$: it is the powerset~$2^\Omega$. Given the
set of events $\events$, a \emph{probability measure} is a function
$\pmeas : \events \rightarrow [0,1]$ such that:
\begin{itemize}
\item $\pmeas(\Omega) = 1$, and 
\item for a collection $E_{i}$ of pairwise disjoint events, $\pmeas(\bigcup_{i}E_{i})=\tensor*[_{\mathbb{R}}]{\sum}{_{i}}\pmeas(E_{i})$,
where $\tensor*[_{\mathbb{R}}]{\sum}{_{i}}a_{i}$
explicitly specifies $a_{i}\in\mathbb{R}$. We may drop the preposing
$\mathbb{R}$ when there is no confusion. 
\end{itemize}

@@

In a strict computational or experimental setting, one may question
the reliance of the definition of probability space on the uncountable
and uncomputable real interval $[0,1]$. This interval includes numbers
like $0.h_1h_2h_3\ldots$ where $h_i$ is 1 or 0 depending on whether
Turing machine $M_i$ halts or not. Such numbers cannot be
computed. This interval also includes numbers like $\frac{\pi}{4}$
which can only be computed with increasingly large resources as the
precision increases.
% \yutsung{Check the meaning of the ``computing $\frac{\pi}{4}$'', because
% Bailey Borwein Plouffe formula can computing the $n$th binary digit of $\pi$ 
% using base 16 directly.}
Therefore, in a resource-aware computational or experimental setting,
it is more appropriate to consider probability measures that map
events to a set of elements computable with a fixed set of
resources. We expand on this observation and then consider
%% two approaches from the literature: set-valued probability
%% measures~\cite{Artstein1972,PuriRalescu1983} and 
interval-valued probability
measures~\cite{Dempster1967,Shafer1976,Weichselberger2000,JamisonLodwick2004}
in detail.\footnote{There is another possible approach that can be
  used to split the real interval $[0,1]$ into a collection of
  subsets~\cite{Artstein1972,PuriRalescu1983} \amr{need to explain the
    connection and why we are not using it.}}

@@

\begin{definition}[Projection Operators;
Orthogonality~\cite{10.2307/2308516,Redhead1987-REDINA,peres1995quantum,Griffiths2003,Swart2013}]
\label{def:Projection} Given a Hilbert space $\Hilb$, an event\footnote{An event
is formally called an experimental proposition~\cite{BirkhoffVonNeumann1936}, a question~\cite{10.2307/2308516,DBLP:journals/corr/abs-0910-2393}, or an
elementary quantum test~\cite{peres1995quantum}.} mathematically is represented as a projection operator $P:\Hilb\rightarrow\Hilb$ onto a linear subspace~$S$ of $\Hilb$. The set of all events can be defined recursively as follow:
\footnote {``Projection'' is sometimes called
``orthogonal projection" or ``self-adjoint projection" to
emphasize $P^{\dagger} = P$~\cite{Maassen2010}.}
\begin{itemize}
\item $\mathbb{0}$ is a projection. 
\item For any pure state~$\ket{\psi}$, $\proj{\psi}$ is a projection
operator. 
\item Projection operators $P_1$ and $P_2$ are orthogonal if $P_1P_2 =
  P_2P_1 = \mathbb{0}$. The sum of two projection operators~$P_{1}\tensor*[_{\mathscr{O}}]{+}{}P_{2}$
is also a projection operator if and only if they are orthogonal, 
where the preposing subscript $\mathscr{O}$
means $\tensor*[_{\mathscr{O}}]{+}{}$ is an operation between operators,
and may be dropped when there is no ambiguity. 
\item Conversely, every projection~$P$ can be expressed as
$\tensor*[_{\mathscr{O}}]{\sum}{_j}\proj{\psi_{j}}$,
where $P$ actually projects onto the linear subspace~$S$ which
has an orthonormal basis~$\{\ket{\psi_{j}}\}$. 
\end{itemize}
\qed\end{definition}

@@

\begin{definition}[Complement; Commutativity;
Independence~\cite{peres1995quantum,Griffiths2003}]~
\begin{itemize}
\item If $P$ is a projection operator, then $\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P$
is also a projection operator, called complement. It is orthogonal to $P$, has
probability~$\mu\left(\mathbb{1}\tensor*[_{\mathscr{O}}]{-}{}P\right)=1\tensor*[_{\mathbb{R}}]{-}{}\mu\left(P\right)$, 
and corresponds to the complement event~$\Omega\backslash E$
in classical probability.
\item Projection operators $P_{1}$ and $P_{2}$ commute if $P_{1}P_{2}=P_{2}P_{1}$.
The product of two projection operators~$P_{1}P_{2}$ is also a projection
operator if and only if they commute. This corresponds to the classical intersection
between event. Given a quantum probability measure $\mu$, two projections
$P_{1}$ and $P_{2}$ are independent if $P_{1}$ and $P_{2}$ commute
and $\mu\left(P_{1}P_{2}\right)=\mu\left(P_{1}\right)\mu\left(P_{2}\right)$.
\end{itemize}
\qed\end{definition}
